---
title: "run_analysis.R ReadMe"
output:
  html_document:
    keep_md: yes
---

I begin my code by cleaning up the global environment and loading the plyr package.  
Note: the Samsung data must be in the home directory.  

```{r}
rm(list=ls())
library(plyr)
```

Next, 
    - I read in the test data and the features and activity information.  
    - Assign the feature descriptions as the column names on the test data.  
    - Using join() from the plyr package, join the activity and test data.  
    - Column bind the activity description and the test data.  
    - Take only the columns expressing: Activity description, Feature mean, Feature standard deviation  
```{r}
# Read in test data and labels
features<-read.table("features.txt")
test<-read.table("X_test.txt")
# Add feature names as column headers
colnames(test)<-features[,2]
# Read in activity labels and data
testLabels<-read.table("y_test.txt")
colnames(testLabels)<-"ActivityCode"
activityLabels<-read.table("activity_labels.txt")
colnames(activityLabels)<-c("ActivityCode","Activity")
# Add acitvity name to training data
activity<-join(testLabels,activityLabels,by="ActivityCode")
test<-cbind(activity$Activity,test)
colnames(test)[1]<-"Activity"
test<-test[,c(1:7,42.47,82:87,122:127,162:167,202,203,215,216,228,229,
              241,242,254,255,267:272,346:351,425:430,504,505,517,518,
              530,531,543,544)]
```

Next,  
    - Use the same procedure on the training data.  
```{r}
# Read in training data and labels
features<-read.table("features.txt")
train<-read.table("X_train.txt")
# Add feature names as column headers
colnames(train)<-features[,2]
# Read in activity labels and data
trainingLabels<-read.table("y_train.txt")
colnames(trainingLabels)<-"ActivityCode"
activityLabels<-read.table("activity_labels.txt")
colnames(activityLabels)<-c("ActivityCode","Activity")
# Add acitvity name to training data
activity<-join(trainingLabels,activityLabels,by="ActivityCode")
train<-cbind(activity$Activity,train)
colnames(train)[1]<-"Activity"
train<-train[,c(1:7,42.47,82:87,122:127,162:167,202,203,215,216,228,229,
              241,242,254,255,267:272,346:351,425:430,504,505,517,518,
              530,531,543,544)]
```
    
Use rbind to combine the two data sets in to a single data set.  
Write the tidy data to a .txt comma seperated file in the home directory.  
Print the str() and summary() of the tidy data set.  
```{r, echo=FALSE}
# Combine the test and trinaing data
data<-rbind(train,test)

write.table(data,"TidyDataSet.txt",sep=",")

print(str(data))
print(summary(data))    
```    
    
Next,  
We will create a second data sets that comprises the means of each Feature by Activity type.  

Start by creating a new data frame with the proper row names and get column names.  

```{r}
#Subsets of data by Activity type

columnNames<-names(train)[2:62]
activities<-c("STANDING","SITTING","LAYING","WALKING",
              "WALKING_DOWNSTAIRS","WALKING_UPSTAIRS")
averages<-data.frame(row.names=activities)
```

Subset the tidy data by Activity description.  
Take the mean of each feature throughout the subsets.  

```{r}
stand<-data[data$Activity=="STANDING",]
sit<-data[data$Activity=="SITTING",]
lay<-data[data$Activity=="LAYING",]
walk<-data[data$Activity=="WALKING",]
downstairs<-data[data$Activity=="WALKING_DOWNSTAIRS",]
upstairs<-data[data$Activity=="WALKING_UPSTAIRS",]

# Data Frame of Averages by Activity and Feature
AvgStand<-colMeans(stand[,-1])
AvgSit<-colMeans(sit[,-1])
AvgLay<-colMeans(lay[,-1])
AvgWalk<-colMeans(walk[,-1])
AvgWalkDownstairs<-colMeans(downstairs[,-1])
AvgWalkUpstairs<-colMeans(upstairs[,-1])
```
   
Create a new data sets by using rbind() to combine the colMeans() results.  
Add the column names.  
Write the new data set to a comma seperated .txt file in the home directory.  
Print the str() and summary() of the new data set.  
   
```{r, echo=FALSE}   
#create a single data frame
averages<-rbind(AvgStand,AvgSit,AvgLay,AvgWalk,AvgWalkDownstairs,AvgWalkUpstairs)
colnames(averages)<-columnNames


write.table(averages,"AvgTidyDataSet.txt",sep=",")

print(str(averages))
print(summary(averages))
```   
   
   
    
    
